A Bayesian Framework for Learning Regular Expressions

### 2012-12-09 paper/paper.tex
Lots of edits.

\section{Results and Discussion}
Figure~\ref{results-chart} shows the comparison between our model and the human subject data.

@@

\caption{Chart showing  $p(s \in \mathcal{L} | S \subseteq {L}) $, sorted in increasing order for model data. Human subject data is in blue, model data is in red. $R$ = .75} 
\label{results-chart}
\end{figure}

@@

Observe that the model's performence is effectively all or none. In twelve of the trials, the model's predictions for  $p(s \in \mathcal{L} | S \subseteq {L}) $ was either 0 or 1. After performing inference, the model generally included one or two hypotheses amounting to over 90\% of the voting weight. Thus, though our inference process considered many hypotheses at once, only a very small number had significant impact on the result.

We note that in cases 16 (S = \{`yyyt', `yt', `yyyt', `yyyt', `yt'\}, s= `yyyyyt'), 17 (S = \{`jojojojo', `jojojojo', `jojojojo'\}, s = `jojo'), and 18 (S = \{`twtwtw', `twtwtw', `twtwtw'\}, s =  `tw'), it is clear that the model more strongly prefers the use of the Kleene star. Here, the prior on the hypothesis is so low that the difference in the likelihood of the base hypothesis as compared to a hypothesis with fewer states and more loops that the posterior probability becomes relatively unlikely. Future work could better determine how this term should be set. However, because this same parameter in part controls for wildcards, lowering it reduces performance on many other cases.

We also observe that in such cases, there was a relatively small set of probable hypotheses, and that any favoring of smaller hypotheses was magnified by this fact.

\section{Conclusion and Future Work}
Despite the number of ways in which our model might be improved mentioned above,  it remains highly implausible that the brain generates its set of candidate hypotheses by ever considering even $2^n$ hypotheses in the first level of beam search we perform, much less more than that. Doing so would be highly inefficient: the number of implausible hypotheses generated would in most cases be far greater than the number of `good' hypotheses. Far more plausible a priori is the hypothesis that the brain uses of number of heuristics to generate a certain set of candidate hypothesis, and then uses some process analogous to Bayesian inference to select among those hypotheses. To shed light on how this hypothesis selection process may occur, we designed an experiment in which candidates chose among a number of possible rules to explain a set of strings S, but have yet to gather enough data to reach any significant conclusions.

Further experiments could be conducted to determine what heuristics may be at work in the hypothesis generation process. For any set of candidate heuristics, it will remain to be determined whether or not such heuristics are at some point learned. Clearly some semantic considerations are at work in human pattern recognition on strings: cat and mat but not qat are meaningful english words; `ao' and `eo' and `io' begin with vowels unlike `xo'. (This experiment was designed to avoid such candidate hypotheses.) One partial solution to this sort of problem would be to permit different varieties of wildcards (e.g., for vowels) beyond just those for number and alphabetical character. Here, though, we risk fitting experimental data at the expense of more general having a more general model, since such wildcards would have to be hardcoded. The model could also be extended to include something like a successor function, which would allow the model to recognize `efg' as a member of the set \{`abc',`bcd',`cde',`def'\}.

We note that we do not directly solve the problem of finding a most plausible way to align a set of strings $S$. Humans are excellent at quickly noticing the pattern in `aall', `ball', and `call'. Although our model will infer that the last three letters can be represented with only three states in each case, in doing so it generates many candidate regular expressions which a human being would not consciously consider.

Future work may be directed towards investigating the hypothesis that humans first align candidate strings and then generate a number of hypotheses. For example, humans would be very quick to notice that ``shrew", ``reward", and ``threw" all contain the `rew' subsequence, and would most likely generate candidate regular expressions which preserved that substring. However, following \citeA{Fernau:2009:ALR:1512988.1513048} we note that the related task of the multiple sequence alignment problem is NP hard, and that a similar model which considers the most likely alignment of a set of strings given many possible alignments will consider a very large number of hypotheses when strings are long enough.  
\bibliographystyle{apacite}
\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}
 \bibliography{regbib}

>>>>

\section{Results and Discussion}
Figure~\ref{results-chart} shows the comparison between our model and the human subject data. In short, there is some agreement ($R$ = 0.78) between human and model data, but there are obvious differences.

@@

\caption{Chart showing  $p(s \in \mathcal{L} | S \subseteq {L}) $, sorted in increasing order for model data. Human subject data is in blue, model data is in red.} 
\label{results-chart}
\end{figure}

@@

It can be observed that the model's performance is effectively all or none; in 13 of the 19 trials, the model's predictions for $p(s \in \mathcal{L} | S \subseteq {L}) $ was within 2\% of either extreme. This is likely caused by the fact that the model generally included one or two hypotheses amounting to over 90\% of the voting weight. Thus, though our inference process considered many hypotheses at once, very few had significant impact on the result.

One potential solution to this all-or-none behavior is to to add noise to the process of accepting a string. For instance, one could allow a small chance of transitioning to an arbitrary state. Alternatively, one could allow the DFA to accept strings ending in a non-accept state with a low probability.

We note that in cases 16 (S = \{`yyyt', `yt', `yyyt', `yyyt', `yt'\}, test = `yyyyyt'), 17 (S = \{`jojojojo', `jojojojo', `jojojojo'\}, test = `jojo'), and 18 (S = \{`twtwtw', `twtwtw', `twtwtw'\}, test =  `tw'), it is clear that the model more strongly prefers the use of the Kleene star. This might suggest that either $\alpha$ or $\beta$ were too high. Future work could better estimate these parameters.

\section{Conclusion and Future Work}
% Despite the potential improvements mentioned above, it remains highly implausible that the brain generates its set of candidate hypotheses by ever considering even $n^2$ hypotheses in the first level of beam search we perform, much less more than that. Doing so would be highly inefficient: the number of implausible hypotheses generated would in most cases be far greater than the number of ``good'' hypotheses. Far more plausible a priori is the hypothesis that the brain uses a number of heuristics to generate a certain set of candidate hypothesis, and then uses some process analogous to Bayesian inference to select among those hypotheses. To shed light on how this hypothesis selection process may occur, we designed an experiment in which candidates chose among a number of possible rules to explain a set of strings $S$, but have yet to gather enough data to reach any significant conclusions.

Further experiments could be conducted to determine what heuristics may be at work in the hypothesis generation process. For any set of candidate heuristics, it will remain to be determined whether or not such heuristics are at some point learned. We note that we do not directly solve the problem of finding a most plausible way to align a set of strings $S$. Humans are excellent at quickly noticing the pattern in `aall', `ball', and `call'. Although our model will infer that the last three letters can be represented with only three states in each case, in doing so it generates many candidate regular expressions which a human being would not consciously consider.

Clearly some semantic considerations are at work in human pattern recognition on strings: cat and mat but not qat are meaningful english words; `ao' and `eo' and `io' begin with vowels unlike `xo'. (This experiment was designed to avoid such candidate hypotheses.) One partial solution to this sort of problem would be to permit different varieties of wildcards (e.g., for vowels) beyond just those for number and alphabetical character. Here, though, we risk fitting experimental data at the expense of having a more general model, since such wildcards would have to be hardcoded. The model could also be extended to include something like a successor function, which would allow the model to recognize `efg' as a member of the set \{`abc',`bcd',`cde',`def'\}.

Further, we might extend the code to use a numeric wildcard. Such a model should be able to infer the structure of phone numbers formatted in a variety of ways: `111-111-1111', `(111) 111 1111', or `111.111.1111'.

Another extension would be to vary the parameters for beam search. We fixed our beam width at 100 for all experiments, but this value could be varied to see if there would be better agreement with subject responses. In particular, and much as in the case of \citeA{vul2009one}, it is possible that we might attain a good model of cognition by assuming an extremely limited search process.

% Future work may be directed towards investigating the hypothesis that humans first align candidate strings and then generate a number of hypotheses. For example, humans would be very quick to notice that ``shrew", ``reward", and ``threw" all contain the `rew' subsequence, and would most likely generate candidate regular expressions which preserved that substring. However, following \citeA{Fernau:2009:ALR:1512988.1513048} we note that the related task of the multiple sequence alignment problem is NP hard, and that a similar model which considers the most likely alignment of a set of strings given many possible alignments will consider a very large number of hypotheses when strings are long enough.
\bibliographystyle{apacite}
\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}
\bibliography{regbib}
